---
title: "EDC Randomization List: Cluster Randomized Trial"
author: "Josh Betz (jbetz@jhu.edu) - Johns Hopkins Biostatistics Center"
date: "`r Sys.time()`"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
bibliography: "cluster_rct.bib"
nocite: '@*'
---

```{r Load-Config-Check-Packages, eval = TRUE, echo = FALSE, message = FALSE}
### User Parameters ############################################################
config_file_name <- "config.yml"
config_name <- "example"

# List of .CSV files - Must contain variables with cluster ID and strata
# Must be located in same directory as .Rmd file
cluster_list_files <-
  c("cluster_list_1.csv",
    "cluster_list_2.csv")


# Need `here` and `config` to load configuration
installed_packages <- installed.packages()[, "Package"]
min_required_packages <- c("here", "config")
if(!all(min_required_packages %in% installed_packages)){
  stop(
    "Required packages must be installed:",
    paste0(
      setdiff(x = min_required_packages, y = installed_packages), 
      collapse = ", "
    )
  )
}

require(here)
 
config_path <-
  file.path(
    here::here(), config_file_name
  )
 
if(file.exists(config_path)) {
  design_config <-
    config::get(
      config = config_name,
      file = config_path
    )
  
  design_config_sha256 <-
    digest::digest(
      object = config_path,
      algo = "sha256",
      file = TRUE
    )
  
  design_config_modtime <-
    file.info(config_path)$mtime
} else {
  stop("Unable to load ", config_file_name, " from ", here::here())
}

required_packages <- design_config$cran_packages
 
if(!all(required_packages %in% installed_packages)) {
  stop(
    "Required packages must be installed:",
    paste0(
      setdiff(x = required_packages, y = installed_packages),
      collapse = ", ")
  )
}

library(digest)
library(dplyr)
library(DT)
library(htmltools)
library(parallel)
library(readr)
library(rlang)
library(stringr)
library(tidyr)

with(
  data = 
    design_config$opts_chunk,
  expr = {
    knitr::opts_chunk$set(
      collapse = collapse,
      echo = echo,
      message = message,
      error = error,
      purl = purl,
      results = results,
      fig.path = fig_path,
      fig.width = fig_width,
      fig.height = fig_height,
      fig.align = fig_align,
      out.width = out_width,
      dpi = dpi
    )
  }
)
```




```{r Extract-Config-Variables}
### Project Information ########################################################
irb_number <-
  design_config$administrative$irb_number
clinicaltrialsgov_nct <-
  design_config$administrative$clinicaltrialsgov_nct
pi_name <-
  design_config$administrative$pi_name
pi_email <-
  design_config$administrative$pi_email %>% 
  paste0("[", ., "](mailto:", ., ")")
statistician_name <-
  design_config$administrative$statistician_name
statistician_email <-
  design_config$administrative$statistician_email %>% 
  paste0("[", ., "](mailto:", ., ")")
edc_pid <-
  design_config$administrative$edc_pid

### Cluster IDs, Treatment Variable ############################################
participant_id_var <- design_config$participant_id_var
cluster_id_var <- design_config$cluster_name_var
cluster_size_var <- design_config$cluster_size_var
clusters_total <- design_config$clusters_total
cluster_tx_var <- design_config$cluster_tx_var
rng_seed <- design_config$rng_parameters$rng_seed

set.seed(
  seed = rng_seed,
  kind = "L'Ecuyer-CMRG"
)

random_seed <- .Random.seed

if(!is.numeric(clusters_total)){
  stop(
    "`clusters_total` (", clusters_total, ") must be numeric."
  )
}




### Stratified Randomization ###################################################
is_stratified <-
  design_config$stratified_randomization

if(is_stratified){
  strata_design <- design_config$strata
  strata_design_vars <- names(strata_design)
  
   randomization_type <-
    paste0(
      "stratified by ",
      paste(
        paste0("`", strata_design_vars, "`"),
        sapply(
          X = strata_design,
          FUN = function(x) 
            paste0("(", paste(paste0("`", x, "`"), collapse = ", "), ")")
        ),
        collapse = "; "
      )
    )
   
} else{
  strata_design_vars <- NULL
  randomization_type <- "without stratification"
}




### Check for Invalid EDC Names ################################################
edc_vars <-
  c(cluster_id_var, cluster_tx_var, strata_design_vars)

invalid_edc_names <-
  stringr::str_detect(
    string = edc_vars,
    pattern = "[^A-Za-z1-9_]"
  )

if(any(invalid_edc_names)){
  stop(
    "Variable names not edc compatible: `",
    paste0(edc_vars[which(invalid_edc_names)], collapse = "`, `"), "`"
  )
}




### Treatment Labels, Block Sizes ##############################################
tx_labels <- design_config$tx_labels
tx_copies_per_block <- design_config$tx_copies_per_block

if(!is.numeric(tx_copies_per_block)){
  stop("`tx_copies_per_block` must be numeric")
}

n_treatments <- length(tx_labels)




### File Name Prefixes #########################################################
file_name_prefix <- design_config$file_name_prefix

invalid_file_prefix <-
  stringr::str_detect(
    string = file_name_prefix,
    pattern = "[^A-Za-z1-9_]"
  )

if(invalid_file_prefix){
  stop("`", file_name_prefix, "` is an invalid file prefix: it must conatin ",
       "only letters, numbers, and underscores.")
}

block_size_description <-
  xfun::join_words(
    words = n_treatments*tx_copies_per_block,
    and = " or "
  )

randomization_type <-
  paste(
    paste(rep(x = 1, times = n_treatments), collapse = ":"),
    "randomization",
    randomization_type
  )

treatment_labels <-
  paste(paste0("`", tx_labels, "`"), collapse = ", ")

inline_config_name <- paste0("`", config_name,"`")
inline_id_var <- paste0("`", cluster_id_var, "`")
inline_tx_var <- paste0("`", cluster_tx_var, "`")
inline_size_var <- paste0("`", cluster_size_var, "`")
```


**NOTE:** Cluster randomized trials require special ethical, statistical, and administrative challenges: See the [references](#references) section for a brief overview of the literature. Choice of a cluster RCT design should be based on considerations of ethics, efficiency, and the research question to be evaluated. Researchers should familiarize themselves with the evolving literature on the design, monitoring, analysis, and reporting of cluster RCTs prior to planning such a study.




## Study Information

Study administration was loaded from the `r inline_config_name` configuration in `r config_file_name`.

  - Configuration: `r config_file_name`
    - Modified: `r design_config_sha256`
    - SHA256 `r design_config_modtime`

information is listed below.

  - IRB Number: `r irb_number`
  - ClinicalTrials.gov Registration: `r clinicaltrialsgov_nct`
  - PI Name: `r pi_name` (`r pi_email`)
  - Statistician Name: `r statistician_name` (`r statistician_email`)
  - EDC PID: `r edc_pid`




### Specification of Randomization:

Pseudorandom number generation is performed using L'Ecuyer's CMRG. Each
randomization list is created using a separate random number stream.
Randomization lists are created by permuted block randomization with random
block sizes.

  - Randomization: `r randomization_type`
    - Treatment Labels: `r treatment_labels`
    - Block sizes: `r block_size_description`
    - Maximum Number of Clusters per Stratum: `r clusters_total`
    - Cluster Identifier in EDC: `r inline_id_var`
    - Treatment Variable Name: `r inline_tx_var`
    - RNG Seed: `r rng_seed`

Lists of sites will have a variable `r inline_size_var`
indicating the maximum number of participants at that cluster.


```{r Check-Lists-Exist}
cluster_lists <- list()

current_list_number <-
  length(cluster_list_files)

cluster_list_file_metadata <-
  data.frame(
    file = as.character(rep(x = NA, times = current_list_number)),
    sha256 = as.character(rep(x = NA, times = current_list_number)),
    modified = as.POSIXct(rep(x = NA, times = current_list_number))
  )

if(any(!file.exists(cluster_list_files))){
  missing_files <-
    cluster_list_files[which(!file.exists(cluster_list_files))]
  stop(
    "Files not found: ", paste0(missing_files, collapse = ", ")
  )
}

for(i in 1:current_list_number) {
  # Store information about each list
  cluster_list_file_metadata$file[i] <- cluster_list_files[i]
  cluster_list_file_metadata$sha256[i] <-
    digest::digest(
      object = cluster_list_files[i],
      algo = "sha256",
      file = TRUE
    )
  cluster_list_file_metadata$modified[i] <-
    file.info(cluster_list_files[i])$mtime

  cluster_lists[[i]] <-
    readr::read_csv(
      file = cluster_list_files[i],
    ) %>%
    dplyr::mutate(
      .cluster_list = i,
      .latest_cluster_list = (i == current_list_number)
    )
}
```




```{r Load-Previous-Output}
rdata_file_name <-
  paste0(file_name_prefix, "_randomization_data_",
         current_list_number, ".Rdata")

# Check for previous output: Load most recent result.
if(current_list_number > 1){
  previous_rdata_files <-
    paste0(file_name_prefix, "_randomization_data_",
           1:(current_list_number - 1), ".Rdata")

  if(!all(file.exists(previous_rdata_files))){
    missing_rdata_files <-
      previous_rdata_files[which(!file.exists(previous_rdata_files))]
    stop(
      "Multiple lists should only be supplied when updating previous output. ",
      "Previous output files were not found: ",
      paste0(missing_rdata_files, collapse = ", ")
    )
  }

  # Load previous result into new environment:
  temp_environment <- new.env()
  
  most_recent_rdata_file <- tail(x = previous_rdata_files, n = 1)
  most_recent_rdata_path <-
    file.path(here::here(), most_recent_rdata_file)

  load(
    file = most_recent_rdata_path,
    envir = temp_environment
  )
  
  rdata_sha256 <-
    digest::digest(
      object = most_recent_rdata_path,
      algo = "sha256",
      file = TRUE
    )
  
  rdata_modtime <-
    file.info(most_recent_rdata_path)$mtime

  import_objects <-
    c(
      "design_config",
      "participant_id_var",
      "all_cluster_lists",
      "all_edc_lists",
      "cluster_list_file_metadata",
      "randomization_list",
      "lists_generated"
    )

  for(i in 1:length(import_objects)){
    assign(
      x = paste0(import_objects[i], "_previous"),
      value =
        get(
          x = import_objects[i],
          envir = temp_environment
        )
    )
  }
  
  previous_data_description <-
    paste0(
      "A randomization list with ", nrow(randomization_list_previous), " rows ",
      "with ", sum(randomization_list_previous$.assigned), " previously ",
      "assigned units."
    )
} else {
  previous_data_description <- "No previous lists have been supplied."
  most_recent_rdata_file <- 
    rdata_sha256 <-
    rdata_modtime <- ""
}
```



  - Previous Data: `r previous_data_description`
    - File: `r most_recent_rdata_file`
    - SHA256: `r rdata_sha256`
    - Modified: `r rdata_modtime`



--------------------------------------------------------------------------------

## Lists of Clusters to Randomize {.tabset}

```{r Print-Metadata}
cluster_list_file_metadata %>% 
  kableExtra::kbl(.) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed")
  )
```



```{r Print-List}
for(i in 1:current_list_number){
  cat("\n### List", i, "\n\n")
  
  cluster_lists[[i]] %>% 
    kableExtra::kbl(.) %>%
    kableExtra::kable_styling(
      bootstrap_options = c("striped", "hover", "condensed")
    ) %>%
    # DT::datatable(
    #   data = .,
    #   extensions = 'Buttons',
    #   options = 
    #     list(
    #       dom = 'Blfrtip',
    #       buttons = c('copy', 'csv', 'excel', 'pdf'),
    #       lengthMenu =
    #         list(
    #           c(10,25,50,-1),
    #           c(10,25,50,"All")
    #         )
    #       )
    # ) %>% 
    # htmltools::tagList() %>% 
    print()
  
  cat("\n\n")
}
```


```{r Check-All-Lists}
for(i in 1:length(cluster_list_files)) {
  cluster_list_i <- cluster_lists[[i]]
  cluster_list_i_cols <- names(cluster_list_i)

  # Check for Expected Variables
  if(!cluster_id_var %in% cluster_list_i_cols){
    stop(
      cluster_list_files[i], " - Cluster ID variable `", cluster_id_var,
      "` not found in data"
    )
  } else if(!cluster_size_var %in% cluster_list_i_cols){
    stop(
      cluster_list_files[i], " - Cluster size variable `", cluster_size_var,
      "` not found in data"
    )
  } else if(is_stratified){
    # If stratified, check for strata variables
    if(!all(strata_design_vars %in% cluster_list_i_cols)){
      missing_vars <-
        setdiff(
          x = strata_design_vars,
          y = cluster_list_i_cols
        )
      stop(
        cluster_list_files[i], " - Strata variables `",
        paste0(missing_vars, collapse = "`, `"), "` not found in data"
      )
    }

    # Check for unexpected values of stratification factors
    for(j in strata_design_vars){
      observed_strata <-
        dplyr::pull(.data = cluster_list_i, var = j) %>%
        unique
      expected_strata <- strata_design[[j]]

      unexpected_strata <-
        setdiff(
          x = observed_strata,
          y = expected_strata
        )

      if(length(unexpected_strata) > 0){
        stop(
          cluster_list_files[i], " - Unexpected values found in stratum `",
          strata_vars[j], "`: `", paste0(unexpected_strata, collapse = "`, `"),
          "`"
        )
      }
    }
  }

  cluster_names <-
    dplyr::pull(.data = cluster_list_i, var = cluster_id_var)

  # Check for invalid ids: null strings, missing values, duplicates
  if(any(cluster_names %in% c("", NA))){
    stop(
      cluster_list_files[i], " - Invalid cluster IDs in rows: ",
      paste0(which(cluster_names %in% c("", NA)), collapse = ", ")
    )
  } else if(any(duplicated(cluster_names))){
    stop(
      cluster_list_files[i], " - Duplicated cluster IDs in rows: ",
      which(duplicated(cluster_names))
    )
  }

  cluster_sizes <-
    dplyr::pull(.data = cluster_list_i, var = cluster_size_var)

  # Check that cluster sizes are not missing, are numeric
  if(!is.numeric(cluster_sizes)){
    stop(
      cluster_list_files[i], " - Non-numeric values in cluster size variable `",
      cluster_size_var, "`"
    )
  } else if(any(is.na(cluster_sizes))){
    stop(
      cluster_list_files[i], " - Invalid cluster sizes rows: ",
      paste0(which(is.na(cluster_sizes)), collapse = ", ")
    )
  }
}
```




```{r Concatenate-Lists}
## What about list 1 has cluster in one strata, list 2 has cluster in another?
all_cluster_lists <-
  do.call(
    what = dplyr::bind_rows,
    args = cluster_lists
  )

repeat_clusters <-
  all_cluster_lists %>%
  dplyr::count(!!sym(cluster_id_var)) %>%
  dplyr::filter(n > 1) %>%
  dplyr::pull(!!sym(cluster_id_var))

# Check for consistency: Cluster listed multiple times with different strata
cluster_strata_check <-
  all_cluster_lists %>%
  dplyr::group_by(
    dplyr::across(
      dplyr::all_of(
        x = c(cluster_id_var, strata_design_vars)
      )
    )
  ) %>%
  dplyr::count() %>%
  dplyr::ungroup() %>%
  dplyr::count(!!sym(cluster_id_var)) %>%
  dplyr::filter(n > 1)

if(nrow(cluster_strata_check) > 0){
  stop(
    "Inconsistent strata for one or more clusters. See `cluster_strata_check`"
  )
}

all_cluster_lists <-
  all_cluster_lists %>%
  dplyr::group_by(
    dplyr::across(
      dplyr::all_of(
        x = c(cluster_id_var)
      )
    )
  ) %>%
  dplyr::mutate(
    `.expand` =
      (dplyr::row_number() > 1)
  ) %>%
  dplyr::ungroup()


### Get Clusters to Expand #####################################################
clusters_to_expand <-
  all_cluster_lists %>%
  dplyr::filter(
    `.latest_cluster_list`,
    `.expand`
  ) %>%
  dplyr::pull(
    !!sym(cluster_id_var)
  )


### Assign Row Number in Stratum-Specific Randomization List ###################
if(is_stratified){
  all_cluster_lists <-
    all_cluster_lists %>%
    dplyr::group_by(
      dplyr::across(
        .col = dplyr::all_of(strata_design_vars)
      )
    )
}


all_cluster_lists <-
  all_cluster_lists %>%
  dplyr::filter(
    `.expand` %in% FALSE
  ) %>%
  dplyr::mutate(
    .cluster_in_stratum = dplyr::row_number()
  ) %>%
  dplyr::ungroup()


clusters_all_names <-
  dplyr::pull(.data = all_cluster_lists, var = cluster_id_var) %>%
  unique

clusters_all_n <- length(clusters_all_names)


### Count Number in Each Stratum: Current List vs. Previous ####################
clusters_by_strata_n <-
  all_cluster_lists %>%
  dplyr::group_by(
    dplyr::across(
      .cols = dplyr::all_of(x = c(".latest_cluster_list", strata_design_vars))
    )
  ) %>%
  dplyr::count() %>%
  dplyr::ungroup()
```




```{r Check-Previous-Output}
if(current_list_number > 1){
  # Check for Consistency with Previous Results
  if(
    !identical(
      x = head(x = cluster_list_file_metadata, -1),
      y = cluster_list_file_metadata_previous
    )
  ){
    stop("File list has changed from previous run. Check chunk",
         "`Check-Previous-Output` for more information.")
  }

  if(!identical(x = design_config, y = design_config_previous)){
    stop("Configuration file `", config_file_name,
         "` has changed from previous run.")
  }

  if(
    !identical(
      x =
      all_cluster_lists %>%
      dplyr::filter(
        `.latest_cluster_list` == FALSE
      ) %>%
      dplyr::select(
        dplyr::all_of(
          x = c(cluster_id_var, strata_design_vars, cluster_size_var)
        )
      ),
      y =
      all_cluster_lists_previous %>%
      dplyr::select(
        dplyr::all_of(
          x = c(cluster_id_var, strata_design_vars, cluster_size_var)
        )
      )
    )
  ){
    stop("List of clusters `all_cluster_lists` has changed from previous run.")
  }
}
```




```{r Create-Randomization-List}
if(current_list_number > 1){
  list_size_check <-
    dplyr::full_join(
      x =
        randomization_list_previous %>%
        dplyr::filter(
          `.assigned` == FALSE
        ) %>%
        dplyr::group_by(
          dplyr::across(
            .cols = dplyr::all_of(x = strata_design_vars)
          )
        ) %>%
        dplyr::count(name = "n_available") %>%
        dplyr::ungroup(),
      y =
        clusters_by_strata_n %>%
        dplyr::filter(
          `.latest_cluster_list` %in% TRUE
        ) %>%
        dplyr::rename(
          n_required = n
        ),
      by = c(strata_design_vars)
    )

  create_new_list <-
    with(data = list_size_check, expr = {any(n_required > n_available)})

  lists_generated <- lists_generated_previous
  
  all_edc_lists <- all_edc_lists_previous

} else if(current_list_number == 1){
  create_new_list <- TRUE

  if(any(clusters_by_strata_n$n > clusters_total)){
    stop(
      "Randomization list contains more cluster IDs (",
      max(clusters_by_strata_n$n), ") than specified in the initial ",
      "configuration file (", clusters_total, ")"
    )
  }

  lists_generated <- 0
  all_edc_lists <- list()
}

if(create_new_list){

  lists_generated <- lists_generated + 1

  # Set a new RNG stream for each list generated
  for(i in 1:lists_generated){
    parallel::nextRNGStream(seed = random_seed)
  }

  if(is_stratified){
    strata_levels <-
      do.call(
        what = tidyr::expand_grid,
        args = design_config$strata
      ) %>%
      dplyr::mutate(
        .stratum = dplyr::row_number()
      )
  } else {
    strata_levels <-
      dplyr::tibble(
        .stratum = 1
      )
  }

  strata_levels_n <- nrow(strata_levels)

  blocks_initial_n <-
    ceiling(clusters_total/(min(tx_copies_per_block)*n_treatments))

  # Random treatment copies per block
  blocks_copies <-
    sample(x = tx_copies_per_block, size = blocks_initial_n, replace = TRUE)

  randomization_list <- list()

  for(i in 1:strata_levels_n){
    # Copy list of treatments to get block size, then permute
    tx_list <-
      sapply(
        X = blocks_copies,
        FUN = function(x, labels = tx_labels)
          sample(x = rep(tx_labels, each = x), replace = FALSE)
      ) %>%
      do.call(
        what = c,
        args = .
      )
    randomization_list[[i]] <-
      data.frame(
        strata_levels[i,],
        `.cluster_in_stratum` =
          1:length(tx_list)
      ) %>%
      dplyr::mutate(
        !!cluster_tx_var := tx_list,
        .assigned = FALSE,
        .randomization_list = lists_generated
      )
  }

  randomization_list <-
    do.call(
      what = dplyr::bind_rows,
      args = randomization_list
    ) 

  # Append to previous list
  if(lists_generated > 1){
    randomization_list <-
      dplyr::bind_rows(
        randomization_list_previous %>% 
          dplyr::mutate(
            `.previous_totals` = `.cluster_in_stratum`
          ) %>% 
          dplyr::mutate(
            .latest_cluster_list =
              dplyr::case_when(
                .latest_cluster_list %in% TRUE ~ FALSE,
              )
          ),
        randomization_list,
      ) %>%
      dplyr::arrange(
        `.stratum`
      ) %>% 
      tidyr::fill(
        `.previous_totals`, 
        .direction = "down"
      ) %>% 
      dplyr::mutate(
        `.cluster_in_stratum` =
          dplyr::case_when(
            .randomization_list == lists_generated ~
              `.previous_totals` + `.cluster_in_stratum`,
            .randomization_list < lists_generated ~
              `.cluster_in_stratum`
          ),
        `.previous_totals` = NULL,
        `.expand` = !!sym(cluster_id_var) %in% clusters_to_expand
      )
  }
}
```




```{r Assign-Treatment}
check_unmatched <-
  dplyr::anti_join(
    x =
      all_cluster_lists %>%
      dplyr::filter(
        .latest_cluster_list == TRUE
      ),
    y = randomization_list,
    by = c(strata_design_vars, ".cluster_in_stratum")
  )

if(nrow(check_unmatched)){
  stop(
    "Clusters unmatched against randomization list - cluster IDs: `",
    paste0(
      dplyr::pull(.data = check_unmatched, var = cluster_id_var),
      collapse = "`, `"
    ), ". See chunk labelled `Assign-Treatment`."
  )
}




if(current_list_number == 1){
  randomization_list <-
    dplyr::left_join(
      x = randomization_list,
      y =
        all_cluster_lists %>%
        dplyr::filter(
          .latest_cluster_list == TRUE
        ),
      by = c(strata_design_vars, ".cluster_in_stratum")
    )
} else {

  clusters_to_assign <-
    dplyr::anti_join(
      x =
        all_cluster_lists %>%
        dplyr::filter(
          .latest_cluster_list %in% TRUE
        ),
      y =
        all_cluster_lists %>%
        dplyr::filter(
          .latest_cluster_list %in% FALSE
        ) %>%
        dplyr::select(
          dplyr::all_of(
            x = c(cluster_id_var)
          )
        ),
      by = c(cluster_id_var)
    )

  randomization_list_new <-
    dplyr::bind_rows(
      # Append Old List
      randomization_list %>%
        dplyr::filter(
          `.assigned` %in% TRUE
        ),

      # Add Newly Assigned Clusters
      dplyr::inner_join(
        x =
          randomization_list %>%
          dplyr::filter(
            `.assigned` %in% FALSE
          ) %>%
          dplyr::select(
            dplyr::all_of(
              x = c(strata_design_vars, ".stratum", ".cluster_in_stratum",
                    cluster_tx_var, ".assigned", ".randomization_list")
            )
          ) %>%
          dplyr::mutate(
            `.assigned` = TRUE
          ),
        y = clusters_to_assign,
        by = c(strata_design_vars, ".cluster_in_stratum")
      ),

      dplyr::anti_join(
        x = randomization_list,
        y =
          all_cluster_lists %>%
          dplyr::select(
            dplyr::all_of(
              x = c(strata_design_vars, ".cluster_in_stratum")
            )
          ),
        by = c(strata_design_vars, ".cluster_in_stratum")
      )
    ) %>%
    dplyr::arrange(
      `.stratum`, `.cluster_in_stratum`
    ) %>%
    dplyr::mutate(
      .latest_cluster_list =
        (.cluster_list == lists_generated) & .assigned
    )

  randomization_list_check <-
    dplyr::anti_join(
      x =
        randomization_list %>%
        dplyr::select(
          dplyr::all_of(
            x = c(strata_design_vars, ".stratum", ".cluster_in_stratum",
                  cluster_tx_var)
          )
        ),
      y =
        randomization_list_new %>%
        dplyr::select(
          dplyr::all_of(
            x = c(strata_design_vars, ".stratum", ".cluster_in_stratum",
                  cluster_tx_var)
          )
        ),
      by = c(strata_design_vars, ".stratum", ".cluster_in_stratum",
             cluster_tx_var)
    )

  if(nrow(randomization_list_check) == 0){
    randomization_list <- randomization_list_new
  } else {
    stop("An error occurred in chunk `Assign-Treatment`")
  }
}
```




--------------------------------------------------------------------------------


## Randomization List

```{r Print-Randomization-List}
randomization_list %>% 
  kableExtra::kbl(
    x = .,
    caption =
      paste0(
        "All randomization lists, concatenated together. Clusters randomized ",
        "from the current list are in **bold font,** while clusters expanded ",
        "(i.e. have additional participant IDs created) are in *italics* with ",
        "red font."
      )
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>% 
  kableExtra::row_spec(
    row = which(randomization_list$.latest_cluster_list),
    bold = TRUE
  ) %>% 
  kableExtra::row_spec(
    row = which(randomization_list$.expand),
    color = "red",
    italic = TRUE
  )
```




```{r Generate-edc-List}
edc_list <-
  randomization_list %>%
  dplyr::filter(
    `.latest_cluster_list` | .expand
  ) %>%
  dplyr::select(
    dplyr::all_of(
      x = c(cluster_id_var, strata_design_vars,
            cluster_tx_var, cluster_size_var)
    )
  ) %>%
  dplyr::group_by(
    dplyr::across(
      .cols = dplyr::all_of(x = c(cluster_id_var))
    )
  ) %>%
  dplyr::mutate(
    .row = paste0(1:cluster_n, collapse = "|")
  ) %>%
  dplyr::ungroup() %>%
  tidyr::separate_longer_delim(
    cols = `.row`,
    delim = "|"
  ) %>%
  dplyr::select(
    dplyr::all_of(x = c(cluster_id_var, cluster_tx_var))
  )

# Mark assigned strata
randomization_list$`.assigned`[
  with(
    data = randomization_list,
    expr = which(`.latest_cluster_list`)
  )
] <- TRUE

all_edc_lists[[current_list_number]] <-
  edc_list
```


```{r Save-Results}
edc_csv_file_name <-
  paste0(file_name_prefix, "_edc_list_", current_list_number, ".csv")

edc_csv_file_path <-
  file.path(here::here(), edc_csv_file_name)

if(file.exists(edc_csv_file_path)){
  stop(
    "File `", edc_csv_file_name, "` exists. This program will not ",
    "overwrite existing files."
  )
} else {
  write.csv(
    x = edc_list,
    file = edc_csv_file_path,
    row.names = FALSE
  )
  
  edc_csv_sha256 <-
    digest::digest(
      object = edc_csv_file_path,
      algo = "sha256",
      file = TRUE
    )
}


tx_list_file_name <-
  paste0(
    file_name_prefix, "_treatment_assignment_", current_list_number, ".csv"
  )

tx_list_file_path <-
  file.path(here::here(), tx_list_file_name)

if(file.exists(tx_list_file_path)){
  stop(
    "File `", tx_list_file_name, "` exists. This program will not ",
    "overwrite existing files."
  )
} else {
  write.csv(
    x = randomization_list,
    file = tx_list_file_path,
    row.names = FALSE
  )
  
  tx_list_sha256 <-
    digest::digest(
      object = tx_list_file_path,
      algo = "sha256",
      file = TRUE
    )
}


rdata_file_name <-
  paste0(file_name_prefix, "_randomization_data_",
         current_list_number, ".Rdata")

rdata_file_path <-
  file.path(here::here(), rdata_file_name)

if(file.exists(rdata_file_path)){
  stop(
    "File `", rdata_file_path, "` exists. This program will not ",
    "overwrite existing files."
  )
} else {
  save(
    list =
      c(
        "design_config",
        "participant_id_var",
        "cluster_id_var",
        "strata_design_vars",
        "cluster_size_var",
        "clusters_total",
        "cluster_tx_var",
        "all_cluster_lists",
        "cluster_list_file_metadata",
        "all_edc_lists",
        "strata_levels",
        "randomization_list",
        "lists_generated"
      ),
    file = rdata_file_path
  )
}

rdata_sha256 <-
  digest::digest(
    object = rdata_file_path,
    algo = "sha256",
    file = TRUE
  )
```




--------------------------------------------------------------------------------


## Results Files

  - Cluster-Level Treatment Assignment:
    - File: `r tx_list_file_name`
    - SHA256: `r tx_list_sha256`
  - Individual-Level Randomization Codes:
    - File: `r edc_csv_file_name`
    - SHA256: `r edc_csv_sha256`
  - Rdata - All Randomization Data:
    - File: `r rdata_file_name`
    - SHA256: `r rdata_sha256`




--------------------------------------------------------------------------------


## References:
