---
title: "Generating Randomization and Treatment Lists for Cluster Randomized Trials"
author: "Josh Betz (jbetz@jhu.edu) - Johns Hopkins Biostatistics Center"
date: "2025-04-21"
output:
  github_document:
    toc: TRUE
bibliography: "cluster_rct.bib"
nocite: '@*'
---

<!-- This is created by README.Rmd -->

```{r Documentation-Setup, echo = FALSE, message = FALSE}
library(dplyr)
library(kableExtra)
library(readr)
```

This repository contains .Rmd reports for generating randomization lists for cluster randomized trials. While it is straightforward to produce a randomization list for a cluster randomized trial, it is not always trivial to link the cluster-level treatment assignment to individuals in those clusters within the [Electronic Data Capture (EDC) system](https://en.wikipedia.org/wiki/Electronic_data_capture).

In an EDC system that (1) allows the use of user-created randomization lists and (2) supports stratified randomization (such as [REDCap](https://projectredcap.org/)), one solution is to "trick" the EDC into treating the clusters as strata in a design with stratified randomization. Rather than having treatment randomly assigned within strata, each cluster's stratum contains only treatment allocations matching the cluster's treatment assignment. This workflow produces such a list that can be uploaded to an EDC, facilitating this approach.

Another solution involves downloading a list of each participant and their corresponding cluster, merging this with the site level treatment assignment, then uploading this merged list to the EDC. This workflow also produces site-level randomization lists. One drawback is that this process may need to be done several times over the course of the study as participants are enrolled.

Other solutions may be possible, such as fields that use SQL joins: before designing a cluster randomized trial, consult the administators of your EDC to determine the best way to carry out such a study.

**This software should be used by a methodologically-trained trialist or statistician. Randomized trials should have ongoing statistical oversight to ensure ethical conduct and scientific validity.**




## How to Use This Repository

#### Make Sure a Cluster Randomized Trial is Appropriate

In an individually randomized trial, treatment is assigned at the level of the individual. In contrast, cluster randomized trials involve randomly allocating treatment to groups of individuals, known as clusters. **Cluster randomized trials require special ethical, statistical, and administrative considerations: See the [references](#references) section for a brief overview of the literature.** Choice of a cluster RCT design should be based on considerations of ethics, efficiency, and the research question to be evaluated. Researchers should familiarize themselves with the evolving literature on the design, monitoring, analysis, and reporting of cluster RCTs prior to planning such a study.




#### 1. Clone the Repository

It is easiest to use this repository with [R](https://cran.r-project.org/) and [R Studio](https://posit.co/download/rstudio-desktop/). Install and update these programs as needed. Update installed packages as needed.

In R Studio, go to `File` > `New Project` > `Version Control` > `Git`: in the Repository URL field, put `https://github.com/jbetz-jhu/cluster_rct_randomization`: choose the subdirectory for the project, and rename the project as you see fit. Click `Create Project` to clone the software from the repository.




#### 2. Install Additional Packages

Next, open `config.yml` in R Studio: this is a configuration file meant to work with the [`config` package in R](https://cran.r-project.org/web/packages/config/index.html). In this file, you'll see a list called `cran_packages` - these are packages that must be downloaded from the [Comprehensive R Archival Network](https://cran.r-project.org/). These can be done using the `Packages` tab in R Studio, or using the command line in R:

```{r install-packages, eval = FALSE}
install.packages(
  c("devtools", "DT", "digest", "dplyr", "here", "htmltools", 
    "kableExtra", "parallel", "readr", "rlang", "stringr", "tidyr",
    "xfun")
```




#### 3. Update the Configuration File

Edit `config.yml` to suit your project. You'll see a heading entitled `example` - This is a configuration for the example files in the repo. You will see another heading entitled `your_study`: modify these entries according to your study requirements. `file_name_prefix` and variable names such as `participant_id_var`, `cluster_name_var`, `cluster_size_var`, `clusters_total`, and `cluster_tx` should be written in [snake case](https://en.wikipedia.org/wiki/Snake_case). `cluster_name_var` should exactly match the name of the variable in the EDC containing the cluster IDs.


#### 4. Create Lists of Clusters to Randomize

The R Markdown files expect the user to supply spreadsheets in [CSV format](https://readr.tidyverse.org/reference/read_delim.html) that specify the clusters to be randomized, and the maximum number of participants anticipated at each cluster. For example, see `cluster_list_1.csv` in the repository:

```{r cluster-list-1, echo = FALSE, message = FALSE, results = "asis"}
readr::read_csv(
  file = "cluster_list_1.csv",
) %>% 
  head() %>% 
  knitr::kable(
    x = .
  )
```

The columns indicate the name of the cluster in the EDC (`cluster_name`) and the maximum number of participants expected at this cluster (`cluster_n`). The values of `cluster_name` should exactly match the values in the EDC. Additionally, two other variables are included: `stratum_1` and `stratum_2` are variables to be used in stratified randomization.

The configuration file `config.yml` tells the R Markdown code how the input files are formatted, how randomization is to be carried out (RNG seeds, stratification factors), as well as metadata about the study.




#### 5. Modify the .Rmd Files

Finally, configure the R Markdown files (.Rmd extensions). Change `config_name` to the name you used in `config.yml` (defaults to `your_study`): 

```{r configuration-name}
config_name <- "your_study"
```

The user specifies `cluster_list_files`, a vector of file names that contain the lists of clusters to be randomized. These files should be in comma-separated value (CSV) format.

```{r example-cluster-list-files, eval = FALSE}
cluster_list_files <-
  c("cluster_list_1.csv")
```

If additional batches of clusters are to be randomized later, their file names should be added in order from first to last.

```{r example-cluster-list-files-2, eval = FALSE}
cluster_list_files <-
  c("cluster_list_1.csv",
    "cluster_list_2.csv")
```



#### 6. Step Through the Code

Go chunk-by-chunk, running code up until the `Save-Results` chunk: built in data checks are meant to identify issues, such as missing information, duplicate cluster IDs within a list, and so on. Hit the `Knit` button to generate a full report.

Note: the code is designed to avoid overwriting files that may contain a previously-created randomization list. Once a list is created, be sure that treatment assignment is backed up in at least two other places. If a list is created for testing, make sure that the file name clearly indicates this in both the files (e.g. set `file_name_prefix = "TESTING_study`) and the treatment labels generated (e.g. `tx_labels` should be `"TESTING_A"` and `"TESTING_B`).




## Adding Additional Individual-Level Allocations

In general, it is advisable to create randomization lists that allow for more clusters and participants than expected. If a cluster has been previously randomized, but additional codes need to be created, create a list with the names (and stratification variables) of previously-randomized strata. These sites should have `TRUE` for the `.expand` column.




## References